### 2016-09-12 draft/draft.tex
Updates

\section{University College London}\label{university-college-london}

\subsubsection{Department of Computer
Science}\label{department-of-computer-science}

\subsection{MSc Summer Project Report}\label{msc-summer-project-report}

\section{QuantiTeam: Blockchain architecture as a medium to verify
collaborative
work}\label{quantiteam-blockchain-architecture-as-a-medium-to-verify-collaborative-work}

\subsection{A Feasibility Study}\label{a-feasibility-study}

\emph{Author:} Benjamin Kremer\\
\emph{Supervisor:} Dr.~Ghita Kouadri Mostefaoui

\emph{This report is submitted as part requirement for the MSc Computer
Science degree at UCL. It is substantially the result of my own work
except where explicitly indicated in the text. The report may be freely
copied and distributed provided the source is explicitly acknowledged.}

September 2016

\begin{abstract}

\end{abstract}

\renewcommand{\abstractname}{Acknowledgements}
\begin{abstract}

\end{abstract}

\setcounter{tocdepth}{3}
% Setting this higher means you get contents entries for
%  more minor section headers.

@@

This project began as an attempt to address this phenomenon of student
disengagement and, over time, developed into a wholly larger
investigation of whether it is possible to utilise blockchain technology
not just to form a network of re-engaged students, but as a tool for the
verification of group-based work at large.\\
The major goals of the project were thus twofold: in a broader sense, to
demonstrate how the properties of a blockchain can be leveraged in a
social, communal context, rather than the financial contexts they are
usually applied to, and more concretely, to create a free, open and
extensible platform for students of all ages to re-engage with their
peers in a productive, educational manner, underpinned by a model of
motivation which rewards collaborative work as a team.

Working from this technological and philosophical objective,
respectively, the following goals were established for the project: -
Create a public Application Programming Interface (henceforth API) to
allow users to interact with a blockchain in a well-defined manner,
independent of a chosen client-side implementation. - Provide an iOS
client-side app for basic task/user/team management to offer a graphical
representation of the API in a format familiar to users. - Establish a
basic deployment of at least 4 distributed validator nodes to implement
a true proof-of-stake consensus
mechanism\cite{BitFuryGroup2015} for the platform.

@@

The following points provide an overview of the author's personal aims
for the project: - Become familiar with native mobile app development by
utilising existing JavaScript knowledge in the context of React
Native\cite{2react-native}. - Learn how
to write smart contracts in Solidity\cite{1solidity} that are useful
in the context of the project and for existing platforms such as
Ethereum\footnote{https://www.ethereum.org/}. - Implement a form of blockchain
technology, by learning how to set up, run \& maintain a a blockchain
proficiently. - Gain a fundamental understanding of Docker\cite{1docker} and how it leverages ``containerisation'' to achieve
straightforward deployments of development environments.

@@

The project followed the methodology of the Unified Process (henceforth
UP) framework for software
development\cite{jacobson1999unified}, and was thus structured in terms of iterative
phases. The system was at first modeled as a whole by using a
theoretical deployment diagram in order to gain an understanding of
which technological entities were required to neatly encapsulate the
responsibilities within the system. Sequence diagrams were then used to
model more complex client-server-blockchain interactions during each UP
iteration, where a new stage usually took the form of defining and
implementing a new domain within the API, e.g.~task management or user
management.

@@

During the implementation stage of the project, a daily log was kept of the key advances and issues encountered during
that day of development. This served as a useful tool to summarise the
progress made, as well as providing an immediate opportunity to document
issues that were encountered.\\
Throughout the implementation stage, required programming work was
broken down into separate units of work by using Github
Issues\footnote{https://developer.github.com/v3/issues/} and organised into a Kanban board\href{}{(REF)}
with the help of waffle.io\footnote{https://waffle.io/}.

As the API continued to grow, unit testing the API's public functions became an
essential part of the project workflow. Due to the primitive error
handling of the still very young Solidity programming language (used to
construct smart contracts for blockchains), any responses coming from
the blockchain had to be rigorously tested server-side in order avoid
silent breakages and regressions.

@@

\paragraph{Chapter 1}\label{chapter-1-1}

@@

\paragraph{Chapter 2}\label{chapter-2}

@@

\paragraph{Chapter 3}\label{chapter-3}

@@

\paragraph{Chapter 4}\label{chapter-4}

@@

\paragraph{Chapter 5}\label{chapter-5}

@@

\paragraph{Chapter 6}\label{chapter-6}

@@

Whereas there is little school staff can do to ensure the attention of a
student within a physical school, supplementing the educational process
with blockchain-based technology may provide a solution to this impasse
between the individual student's aims and that of the educator. By
establishing a system in which collaborative work is directly linked to
both an individual and social reward, students are able to focus on a
more immediate \& manageable target than an abstract and potentially
far-off seeming concept of a high school diploma.

@@

\paragraph{SocialX}\label{socialx}

@@

\paragraph{WikiSpaces}\label{wikispaces}

@@

\paragraph{HaikuLearning}\label{haikulearning}

HaikuLearning\cite{haikulearning}, similarly to WikiSpaces, is a web app which attempts to
emulate and enhance the classroom experience for both school teachers
and students. It offer a unified environment to hand in assignments and
to receive feedback \& grades, as well as enabling integrations with
Google Apps for extensibility.

@@

Before proceeding with the following section, two frequently used
concepts should be clearly delineated: blockchains and smart contracts.
A blockchain is a data structure composed of a list of blocks of
transactions, where each block is linked back to the previous block in
the chain\cite{antonopoulos2014mastering}. Each block is identified by a
hash, which is stored in the block's header alongside the previous
block's hash, known as the parent block\cite{antonopoulos2014mastering}, thus providing a
traceable, reverse-chronological chain of transactions over time.\\
The concept of a smart contract was originally developed by Nick Szabo
in 1997, and is usually evoked in terms of a digital protocol which
enforces and/or verifies the performance of a contract between two or
more
parties\cite{1szabo}.

At the outset of the project, a key decision which had to be made was
the choice of blockchain implementation to be used. This decision was
critical, as discovering a major flaw in the implementation or an
incompatibility between what the implementation could provide and what
was needed, would have resulted in (at least temporary) deadlock for the
entire project.\\
The author's initial experimentation with
OpenChain\footnote{https://www.openchain.org/}, which seemed to
offer the possibly useful ability to chain contracts to each other, was
quickly abandoned as the available documentation was insufficiently
detailed and partially ambiguous. This would have left the project open
to a lot of guesswork and thus the above mentioned risk of deadlock.\\
MultiChain\footnote{http://www.multichain.com/}, on the other
hand, was an implementation which had well-defined documentation, but it
assumed a level of pre-existing familiarity with the API of a
blockchain, as well as providing preciously little context as to how
smart contracts could be developed for the platform in an effective
way.\\
The apparent drawbacks of the two mentioned blockchain implementations
therefore led the author to the
Eris\cite{1erisindustries} platform. Eris itself
provides a suite of tools which wrap and augment a
Tendermint\cite{1tendermint} blockchain
implementation. Eris's wealth of documentation and its well-structured
command line interface (henceforth CLI) were the deciding factors for it
becoming the blockchain implementation of choice. The documentation
provided not only step-by-step examples of how a developer could
configure and deploy a blockchain, but also provided in-depth tutorials
and examples on how to write \& deploy
Solidity\cite{1solidity}
smart contracts for the platform; highly valuable for a smart contract
novice such as the author. Furthermore, the fact that Tendermint is an
entirely separate software project meant that there was an additional
repository of documentation for this specific type of blockchain and the
computer science theory backing it.

@@

\paragraph{Express}\label{express}

@@

\paragraph{eris-wrapper \& eris-logger}\label{eris-wrapper-eris-logger}

The eris-wrapper \& eris-logger modules were adopted from a ``Hello
World''-style example Eris provides to showcase how a NodeJS server
implementing their platform could be
structured\footnote{https://github.com/eris-ltd/hello-eris}.\\
eris-wrapper provides a convenient abstraction of low-level bindings
that have to take place between the NodeJS server and the Tendermint
blockchain, while eris-logger simply sets up a simple logger with
different logging types, such as \texttt{ERROR}, \texttt{INFO} and
\texttt{DEBUG}.

\paragraph{Async}\label{async}

@@

Based on these two lessons learnt, the author decided to use Facebook's
React Native\cite{2react-native} for the project, which grafts a truly native mobile development
framework on top of the outstanding React library, along with the now
familiar Redux library.\\
React Native offers all the benefits of React (high levels of modularity
\& encapsulation, composability) while allowing the developer to hook
into events and effects of the native mobile operating system. This
means that transitioning between views smoothly or sending a push
notification, for example, becomes a trivial undertaking.\\
Redux provides a well-defined interface to manage the state of the
application's data by managing all of it through a centralised
\texttt{store} object, the contents of which can be altered exclusively
by a set of actions previously defined by the developer. This provides a
level of clarity of how the application's state mutates over time that
is hard to achieve in a traditional Model-View-Controller\cite{krasner1988description}
(henceforth MVC) approach, in which many controllers have access to the
application's state simultaneously.

@@

\paragraph{redux-thunk}\label{redux-thunk}

@@

\paragraph{redux-logger}\label{redux-logger}

@@

\paragraph{tcomb-form-native}\label{tcomb-form-native}

@@

Composing unit tests for each public function of the system's API was an
essential part of the development process, especially as the Tendermint
blockchain was able to report errors within its own processes to the
NodeJS server only in the most rudimentary of terms. This meant that if
no testing suite was in place, silent breakages and undefined behaviour
were likely to occur as the API grew.

@@

\paragraph{Mocha}\label{mocha}

@@

\paragraph{Chai}\label{chai}

Chai\footnote{http://chaijs.com/} is an assertion library for both
test-driven development (TDD) and behaviour-driven development (BDD).
More specifically, Chai's TDD \texttt{assert} function was used to keep
test assertions simple and testable against an expected result.

\paragraph{Istanbul}\label{istanbul}

@@

GanttPro\footnote{https://ganttpro.com/}, an online creator and
editor for Gantt charts, was used to create a timeline for the project,
while Google Docs\footnote{https://www.google.com/docs/about/}
in conjunction with draw.io\footnote{https://www.draw.io/} were
used to capture requirements and create UML diagrams.\\
This report was originally written in markdown and transposed to the
Latex format using pandoc\footnote{http://pandoc.org/}.

@@

\paragraph{Editor}\label{editor}

@@

\paragraph{Static Type Analyser}\label{static-type-analyser}

Flowtype\cite{1flowtype}, a static type
analyser for JavaScript, is a further in-built feature of the Nuclide
plugin. Flowtype allows the developer to define a type for a variable, a
type signature for a function, and even to create custom union \&
intersection types. Flowtype then checks whether the defined type
specifications are adhered to and warns if it detects a TypeError. This
is an incredibly useful tool for a dynamically-typed language such as
JavaScript, where unintended type coercion is a common ailment.

\paragraph{Debugging}\label{debugging}

@@

\paragraph{Version Control}\label{version-control}

@@

\paragraph{Docker \& Shell Scripts}\label{docker-shell-scripts}

@@

As outlined in chapters 1 \& 2, the key problems the project was trying
to solve the way in which students relate to their school work on an
everyday basis, as well as investigating the feasibility of a system
which uses a distributed blockchain data structure to verify and store
data about events which represent collaborative work.\\
The conceptual diagram below shows how the basic components and
interactions within this system might look.

\begin{figure}[htbp]
\centering
\includegraphics[width=0.7\textwidth]{../diagrams/conceptDiagram.png}
\caption{An early conceptual diagram demonstrating possible interactions
with the system}
\end{figure}

@@

\include{functional-requirements}

\include{nonfunctional-requirements}

@@

\paragraph{Factory Contracts}\label{factory-contracts}

@@

\paragraph{Data Structure Contract}\label{data-structure-contract}

@@

\paragraph{Manager Contracts}\label{manager-contracts}

@@

\paragraph{Linker Contract}\label{linker-contract}

@@

\paragraph{Action Events}\label{action-events}

Solidity's lack of any real logging capabilities meant there needed to
be consideration in regard to how function calls within contracts could
be observed from the outside. Solidity does, fortunately, provide an
\texttt{event} type, which opened up the possibility of triggering such
an \texttt{event} where required and listening for it server-side, where
logging the events was a trivial matter.

@@

\paragraph{API Router}\label{api-router}

Once the structure of the smart contracts had been established, the
analysis of the server-side implementation quickly revealed that the
server's structure would largely mirror that of the smart contracts.
This was the case partially due to the way Eris's JavaScript library was
implemented, but largely due to the fact that this would keep the the
final API succinct and free of unnecessary cognitive load for the author
and for any future developers.\\
The server would therefore simply act as a relay and transformer for
data travelling between the client-side application and the blockchain,
acting as a \emph{de facto} middleman. In more concrete terms, this
would involve calling relevant contract methods on the blockchain when a
certain API endpoint was requested and transforming data between
hexadecimal and UTF-8, for example. This is necessary due to Solidity's
poor support for strings at the time of writing, meaning that all
strings would have to be encoded into 32-byte fields of hexadecimal,
known in Solidity as the \texttt{bytes32} type.

\paragraph{Uploader}\label{uploader}

@@

\paragraph{Abstracting view components with
React}\label{abstracting-view-components-with-react}

React takes markedly different approach to view templating compared to
the other major web development frameworks such as Google's
AngularJS\footnote{https://angularjs.org/}. React's philosophy
focuses on breaking a view down into its constituent parts to form
reusable components. These components can then be composed in whichever
way the developer sees fit. This level of composability comes to its
full potential in React Native, as it enables the abstraction of common
UI components within a mobile app, providing ample opportunity to even
reuse said components across different platforms. The decision to
abstract a given component was therefore made according to the following
criteria, ordered by highest weighting first:

@@

\emph{Cross-platform potential} - A secondary consideration was whether
the abstraction would be useful across platforms, i.e.~between iOS and
Android. For the initial iteration of QuantiTeam which would only focus
on iOS this aspect was less significant than the frequency of use, but
it played a significant role in keeping the client-side application open
to extension at a later point nonetheless.

\paragraph{Managing state with Redux}\label{managing-state-with-redux}

While considering the responsibilities that a client-side implementation
would have to fulfill to meet the established requirements, it also
became clear that even for a simple implementation there was a
considerable amount of application state that would have to be managed
by the client. Besides being the most common choice within React Native
applications, the Redux library provides a well-structured approach to
state management, derived from the philosophy behind the Elm programming
language. Redux ensures that events which mutate state are well-defined
and that they may only take place in a single direction
(unidirectionally).\\
In concrete terms, Redux achieves these orderly state mutations by
following three interdependent
principles\cite{1redux}:

@@

QuantiTeam broadly follows the three-tier architecture of a typical
Model-View-Controller\cite{krasner1988description} application with the important distinction that the
roles within the MVC pattern are applied to an entire system of various
applications, rather than a single application. In concrete terms, this
means that blockchain represents the Model element by establishing the
system's data model through the smart contracts applied to it, while the
NodeJS server represents the Controller element, providing a public
interface for client applications to issue requests to the blockchain
and handling raw responses from the blockchain. The ``View'' element of
the implementation is therefore interchangeable, as the RESTful API
formed by the web server and the blockchain provides a uniform set of
endpoints to communicate with, tying no special or unique value to the
client, in this case a mobile app.

@@

In order to mirror the described approach of creating a standalone unit
out of the Model and Controller aspects of the system's architecture,
the project's source code was structured into two directories at the
highest level. The \texttt{app} directory contains the React Native
client-side application, while the \texttt{chain} directory holds
configurations and environment variables to run the Tendermint
blockchain via the \texttt{eris} CLI, and provides the REST API server
in its \texttt{js} subdirectory.

@@

\paragraph{Strings and Numbers}\label{strings-and-numbers}

@@

\paragraph{Arrays and Objects}\label{arrays-and-objects}

@@

\paragraph{Data: Factory Contracts}\label{data-factory-contracts}

@@

\paragraph{Operations: Manager
Contracts}\label{operations-manager-contracts}

@@

\paragraph{Relations: Linker Contract}\label{relations-linker-contract}

@@

REST, which is an acronym for Representational State Transfer, is a
commonly used web development pattern which attempts to ensure
reliability and scalability of the web service implementing
it\cite{1rest}. Within
the context of QuantiTeam, achieving RESTfulness was key for the
system's API, as this would help ensure the system's usefulness to any
context of client-side implementation, rather than specifically a mobile
paradigm.\\
This subsection therefore describes the properties of a RESTful service
and how each applies to QuantiTeam's server-side architecture.

\paragraph{Client-Server Dichotomy}\label{client-server-dichotomy}

@@

\paragraph{Stateless}\label{stateless}

@@

\paragraph{Cacheable}\label{cacheable}

@@

\paragraph{Uniform Interface}\label{uniform-interface}

@@

\paragraph{Layered System}\label{layered-system}

@@

Within the MVC pattern, views are the component responsible for the
graphical representation of the application's -- or in this case the
system's -- data
models\cite{krasner1988description}.
Although a view may also be described as ``a visual representation of
models that present a filtered view of the current
state''\href{https://addyosmani.com/resources/essentialjsdesignpatterns/book/\#detailmvc}{(Learning
JS Design Patterns)}, the parallels between the MVC pattern and
QuantiTeam's structure become somewhat less applicable. While the
client-side React Native app does of course act as a filtered graphical
representation of the blockchain's models, it also contains its own
local state and therefore deviates from the typical description of an
MVC pattern's view.

@@

The snippet above shows how Flowtype enabled the author to define what
type primitives a \texttt{User} object's fields should adhere to. This
significantly reduced bug frequency, both within the app and the API
itself, by catching inadvertent type coercions or possibly undefined
values before the app's runtime environment was even entered.\\
Furthermore, the \texttt{User} type exemplifies how Flowtype allows a
differentiation between mandatory (\texttt{username}) and optional
(\texttt{teamname?}) fields, adding an element of flexibility where
needed. For example, the \texttt{address?} field cannot be defined at
the time a \texttt{User}-type object is instantiated in the client-side
application, as the hexadecimal address for the User contract can only
be returned to the client once the user has been registered in the
blockchain post-instantiation.

Finally, Flowtype elevated the ability to define actions and the
expected types of their payloads in a more fine-grained manner compared
to regular JavaScript, by enabling the author to add type constraints to
variables and functions where needed.

@@

\paragraph{Frequency \& Variability of
Use}\label{frequency-variability-of-use}

@@

\paragraph{Cross-Platform Potential}\label{cross-platform-potential}

@@

\paragraph{Synchronous UI Actions}\label{synchronous-ui-actions}

@@

\paragraph{Asynchronous I/O Actions}\label{asynchronous-io-actions}

@@

Once the required communication abilities for the system had been
established, meaning the author was able to trigger smart contract
events within the blockchain via HTTP requests, the system's API was
established and expanded via test-driven development. This meant that
any new functions of the system would be developed in the following
manner\cite{beck2003test}:

@@

=============================== Coverage summary ===============================
Statements   : 91.28% ( 429/470 ), 7 ignored
Branches     : 58.02% ( 76/131 ), 7 ignored
Functions    : 98.56% ( 137/139 ), 3 ignored
Lines        : 91.86% ( 429/467 )
================================================================================
\end{lstlisting}

@@

\paragraph{Verifiability}\label{verifiability}

>>>>

\include{frontmatter}

\include{abstract}

\renewcommand{\abstractname}{Acknowledgements}
\begin{abstract}
First and foremost, I would like to thank my supervisor, Dr. Ghita Kouadri Mostefaoui for her continuous guidance throughout this project. I would also like to thank my family, my flatmates, and Kimeshan Naidoo \& Aditya Mukherjee in particular for their continuous support.

I owe a special thank you to Zach Ramsay from Eris Industries, whose enthusiasm for what I was trying to achieve provided a wealth of motivation.
\end{abstract}

\setcounter{tocdepth}{2}
% Setting this higher means you get contents entries for
%  more minor section headers.

@@

This project began as an attempt to address this phenomenon of student
disengagement and, over time, developed into a larger
investigation of whether it is possible to utilise blockchain technology
not just to form a network of re-engaged students, but as a tool for the
verification and quantification of group-based work at large.\\
The major motivations that guided the project were thus twofold: in a broader sense, to
demonstrate how the properties of a blockchain can be leveraged in a
social, communal context, rather than the financial contexts they are
usually applied to, and more concretely, to create a free, open and
extensible platform for students of all ages to re-engage with their
peers in a productive, educational manner, underpinned by a model of
motivation which rewards collaborative work as a team.

Working from this technological and philosophical motivation,
respectively, the following goals were established for the project:
\begin{itemize}
\item Create a public Application Programming Interface (henceforth API) to allow users to
interact with a blockchain in a well-defined manner, independent of a chosen client-side implementation.
\item Provide an iOS client-side app for basic task, user, and team management to offer a graphical
representation of the API in a format familiar to non-technical individuals.
\end{itemize}

@@

The following points provide an overview of the author's personal aims
for the project:
\begin{itemize}
\item Become familiar with native mobile app development by
utilising existing JavaScript knowledge in the context of React Native\cite{2react-native}.
\item Learn how to write smart contracts in Solidity\cite{1solidity} that are useful
in the context of the project and for existing platforms such as Ethereum\footnote{https://www.ethereum.org/}.
\item Implement a form of blockchain technology, by learning how to set up, run \& maintain a blockchain proficiently.
\item Gain a fundamental understanding of Docker\cite{1docker} and how it leverages ``containerisation'' to achieve straightforward deployments of development environments.
\end{itemize}

@@

The project followed the methodology of the Unified Process\cite{jacobson1999unified} (henceforth
UP) framework for software
development, and was thus structured in terms of iterative
phases. The system was at first modeled as a whole by using a
theoretical deployment diagram in order to gain an understanding of
which technological entities were required to neatly encapsulate the
responsibilities within the system. Sequence diagrams were then used to
model more complex client-server-blockchain interactions during each UP
iteration, where a new stage usually took the form of defining and
implementing a new domain within the API, e.g.~task management or user
management.

@@

During the implementation stage of the project, a daily log was kept of the key advances and issues encountered during
that day of development. This served as a useful tool to summarise the
progress made, as well as providing an immediate opportunity to document
issues that were encountered.\\
Throughout the implementation stage, required programming work was
broken down into separate units of work by using Github
Issues\footnote{https://developer.github.com/v3/issues/} and organised into a Kanban\cite{1kanban} board with the help of waffle.io\footnote{https://waffle.io/}.

As the API continued to grow, unit testing the API's public functions became an
essential part of the project workflow. Due to the primitive error
handling of the still very young Solidity programming language (used to
construct smart contracts for the blockchain), any responses coming from
the blockchain had to be rigorously tested server-side in order avoid
silent breakages and regressions.

@@

\subsubsection{Chapter 1}\label{chapter-1-1}

@@

\subsubsection{Chapter 2}\label{chapter-2}

@@

\subsubsection{Chapter 3}\label{chapter-3}

@@

\subsubsection{Chapter 4}\label{chapter-4}

@@

\subsubsection{Chapter 5}\label{chapter-5}

@@

\subsubsection{Chapter 6}\label{chapter-6}

@@

Whereas there is little school staff can do to ensure the attention of a
student within a physical school, supplementing the educational process
with blockchain-based technology may provide a solution to this impasse
between the individual student's aims and that of the educator. By
establishing a system in which collaborative work is directly linked to
both an individual and social reward, students are able to focus on a
more immediate \& manageable target than an abstract and potentially
far-off seeming concept of, for example, a high school diploma.

@@

\subsubsection{SocialX}\label{socialx}

@@

\subsubsection{WikiSpaces}\label{wikispaces}

@@

\subsubsection{HaikuLearning}\label{haikulearning}

HaikuLearning\cite{haikulearning}, similarly to WikiSpaces, is a web app which attempts to
emulate and enhance the classroom experience for both school teachers
and students. It offers a unified environment to hand in assignments and
to receive feedback \& grades, as well as enabling integrations with
Google Apps for extensibility.

@@

Before proceeding with the following section, two frequently used
concepts should be clearly delineated: blockchains and smart contracts.
A blockchain is a data structure composed of a list of grouped transactions, known as blocks, where each block is linked back to the previous block in
the chain\cite{antonopoulos2014mastering}. Each block is identified by a
hash, which is stored in the block's header alongside the previous
block's hash, known as the parent block\cite{antonopoulos2014mastering}, thus providing a
traceable, reverse-chronological chain of transactions over time.\\
The concept of a smart contract was originally developed by Nick Szabo
in 1997, and is usually evoked in terms of a digital protocol which
enforces and/or verifies the performance of a contract between two or
more
parties\cite{1szabo}.

At the outset of the project, a key decision which had to be made was
the choice of blockchain implementation to be used. This decision was
critical, as discovering a major flaw in the implementation or an
incompatibility between what the implementation could provide and what
was needed, would have resulted in (at least temporary) deadlock for the
entire project.\\
The author's initial experimentation with
OpenChain\footnote{https://www.openchain.org/}, which seemed to
offer the possibly useful ability to chain contracts to each other, was
quickly abandoned as the available documentation was insufficiently
detailed and partially ambiguous. This would have left the project open
to a lot of guesswork and thus the above mentioned risk of deadlock.\\
MultiChain\footnote{http://www.multichain.com/}, on the other
hand, was an implementation which had well-defined documentation, but it
assumed a level of pre-existing familiarity with the API of a
blockchain, as well as providing preciously little context as to how
smart contracts could be developed for the platform in an effective
way.\\
The apparent drawbacks of the two mentioned blockchain implementations
therefore led the author to the
Eris\cite{1erisindustries} platform. Eris itself
provides a suite of tools which wrap and augment a
Tendermint\cite{1tendermint} blockchain
implementation. Eris's wealth of documentation and its well-structured
command line interface (henceforth CLI) were the deciding factors for it
becoming the blockchain implementation of choice. The documentation
provided not only step-by-step examples of how a developer could
configure and deploy a blockchain, but also provided in-depth tutorials
and examples on how to write \& deploy
Solidity\cite{1solidity}
smart contracts for the platform; highly valuable for a smart contract
novice such as the author. Furthermore, the fact that Tendermint is an
entirely separate software project meant that there was an additional
repository of documentation for this specific type of blockchain and the computer science it is based on.

@@

\subsubsection{Express}\label{express}

@@

\subsubsection{eris-wrapper \& eris-logger}\label{eris-wrapper-eris-logger}

The eris-wrapper \& eris-logger modules were adopted from a ``Hello
World''-style example Eris provides to showcase how a NodeJS server
implementing their platform could be
structured\footnote{https://github.com/eris-ltd/hello-eris}.\\
eris-wrapper provides a convenient abstraction of low-level bindings
that have to take place between the NodeJS server and the Tendermint
blockchain, while eris-logger simplifies setting up a logger with
different logging types, such as \texttt{ERROR}, \texttt{INFO} and
\texttt{DEBUG}.

\subsubsection{Async}\label{async}

@@

Based on these two lessons learnt, the author decided to use Facebook's
React Native\cite{2react-native} for the project, which grafts a native mobile development
framework on top of the outstanding React library, along with the now
familiar Redux library.\\
React Native offers all the benefits of React while allowing the developer to hook
into events and effects of the native mobile operating system. This
means that transitioning between views smoothly or sending a push
notification, for example, becomes a trivial undertaking.\\
Redux provides a well-defined interface to manage the state of the
application's data by managing all of it through a centralised
\texttt{store} object, the contents of which can be altered exclusively
by a set of actions previously defined by the developer. This provides a
level of clarity of how the application's state mutates over time that
is hard to achieve in a traditional Model-View-Controller\cite{krasner1988description}
(henceforth MVC) approach, in which many controllers have access to the
application's state simultaneously.

@@

\subsubsection{redux-thunk}\label{redux-thunk}

@@

\subsubsection{redux-logger}\label{redux-logger}

@@

\subsubsection{tcomb-form-native}\label{tcomb-form-native}

@@

Composing unit tests for each public function of the system's API was an
essential part of the development process, especially as the Tendermint
blockchain was able to report errors within its own processes only in the most rudimentary of terms. This meant that if
no testing suite was in place, silent breakages and undefined behaviour
were likely to occur as the API grew.

@@

\subsubsection{Mocha}\label{mocha}

@@

\subsubsection{Chai}\label{chai}

Chai\footnote{http://chaijs.com/} is an assertion library for both
test-driven development (TDD) and behaviour-driven development (BDD).
Chai's TDD \texttt{assert} function was used to keep
test assertions simple and testable against an expected result.

\subsubsection{Istanbul}\label{istanbul}

@@

GanttPro\footnote{https://ganttpro.com/}, an online creator and
editor for Gantt charts, was used to create a timeline for the project,
while Google Docs\footnote{https://www.google.com/docs/about/}
in conjunction with draw.io\footnote{https://www.draw.io/} were
used to capture requirements and create UML diagrams.\\
This report was originally written in Markdown and transposed to the
Latex format using pandoc\footnote{http://pandoc.org/}.

@@

\subsubsection{Editor}\label{editor}

@@

\subsubsection{Static Type Analyser}\label{static-type-analyser}

Flowtype\cite{1flowtype}, a static type
analyser for JavaScript, is a further in-built feature of the Nuclide
plugin. Flowtype allows the developer to define a type for a variable, a
type signature for a function, and even to create custom union \&
intersection types. Flowtype then checks whether the defined type
specifications are adhered to and warns if it detects a TypeError. This
is an incredibly useful tool for a dynamically-typed language such as
JavaScript, where unintended type coercion is a common issue.

\subsubsection{Debugging}\label{debugging}

@@

\subsubsection{Version Control}\label{version-control}

@@

\subsubsection{Docker \& Shell Scripts}\label{docker-shell-scripts}

@@

\begin{figure}[h]
\centering
\includegraphics[width=0.7\textwidth]{../diagrams/conceptDiagram.png}
\caption{An early conceptual diagram demonstrating possible interactions
with the system}
\end{figure}

As outlined in chapters 1 \& 2, the key problem the project attempts
to solve is the way in which students relate to their school work on an
everyday basis by investigating the feasibility of a system
which uses a distributed blockchain data structure to verify and store
data about events which represent collaborative work.\\
Figure 3.1 shows how the basic components and
interactions within this system might look.

@@

\include{functional-requirements}
\include{nonfunctional-requirements}

@@

\subsubsection{Factory Contracts}\label{factory-contracts}

@@

\subsubsection{Data Structure Contract}\label{data-structure-contract}

@@

\subsubsection{Manager Contracts}\label{manager-contracts}

@@

\subsubsection{Linker Contract}\label{linker-contract}

@@

@@

\subsubsection{API Router}\label{api-router}

Once the structure of the smart contracts had been established, the
analysis of the server-side implementation quickly revealed that the
server's structure would largely mirror that of the smart contracts.
This was the case partially due to the way Eris's JavaScript library was
implemented, but largely due to the fact that this would keep the the
final API succinct and free of unnecessary cognitive load for the author
and for any future developers.\\
The server would therefore simply act as a relay and transformer for
data travelling between the client-side application and the blockchain,
acting as a \emph{de facto} middleman. In more concrete terms, this
would involve calling relevant contract methods on the blockchain when a
certain API endpoint was requested and transforming data between
hexadecimal and UTF-8 encoding, for example. This is necessary due to Solidity's
poor support for strings at the time of writing, meaning that all
strings would have to be encoded into 32-byte fields of hexadecimal,
known in Solidity as the \texttt{bytes32} type.

\subsubsection{Uploader}\label{uploader}

@@

\subsubsection{Abstracting view components with
React}\label{abstracting-view-components-with-react}

React takes markedly different approach to view templating compared to
the other major web development frameworks such as Google's
AngularJS\footnote{https://angularjs.org/}. React's philosophy
focuses on breaking a view down into its constituent parts to form
reusable components. These components can then be composed in whichever
way the developer sees fit. This level of composability comes to its
full potential in React Native, as it enables the abstraction of common
UI components within a mobile app, providing ample opportunity to reuse said components even across different platforms. The decision to
abstract a given component was therefore made according to the following
criteria, ordered by highest weighting first:

@@

\emph{Cross-platform potential} - A secondary consideration was whether
the abstraction would be useful across platforms, i.e.~between iOS and
Android. For the initial iteration of QuantiTeam, which would only focus
on iOS, this aspect was less significant than the frequency of use, but
it played a significant role in keeping the client-side application open
to extension at a later point nonetheless.

\subsubsection{Managing state with Redux}\label{managing-state-with-redux}

While considering the responsibilities that a client-side implementation
would have to fulfill to meet the established requirements, it also
became clear that even for a simple implementation there was a
considerable amount of application state that would have to be managed
by the client. Besides being the most common choice within React Native
applications, the Redux library provides a well-structured approach to
state management, derived from the philosophy behind the Elm\footnote{http://elm-lang.org/} programming
language. Redux ensures that events which mutate state are well-defined
and that they may only take place in a single direction
(unidirectionally).\\
In concrete terms, Redux achieves these orderly state mutations by
following three interdependent
principles\cite{1redux}:

@@

@@

QuantiTeam broadly follows the three-tier architecture of a typical
Model-View-Controller\cite{krasner1988description} application with the important distinction that the
roles within the MVC pattern are applied to an entire system of various
applications, rather than a single application. In concrete terms, this
means that blockchain represents the Model element by establishing the
system's data model through the smart contracts applied to it, while the
NodeJS server represents the Controller element, providing a public
interface for client applications to issue requests to the blockchain
and handling raw responses from the blockchain. The View element of
the implementation is therefore interchangeable, as the RESTful API
formed by the web server and the blockchain provides a uniform set of
endpoints to communicate with, tying no special or unique value to the
client, in this case a mobile app.

@@

\subsubsection{Strings and Numbers}\label{strings-and-numbers}

@@

\subsubsection{Arrays and Objects}\label{arrays-and-objects}

@@

\subsubsection{Data: Factory Contracts}\label{data-factory-contracts}

@@

\subsubsection{Operations: Manager
Contracts}\label{operations-manager-contracts}

@@

\subsubsection{Relations: Linker Contract}\label{relations-linker-contract}

@@

REST, which is an acronym for Representational State Transfer, is a
commonly used web development pattern which attempts to ensure
reliability and scalability for the web service implementing
it\cite{1rest}. Within
the context of QuantiTeam, achieving RESTfulness was key for the
system's API, as this would help ensure the system's usefulness to any
context of client-side implementation, rather than specifically a mobile
paradigm.\\
This subsection therefore describes the properties of a RESTful service
and how each applies to QuantiTeam's server-side architecture.

\subsubsection{Client-Server Dichotomy}\label{client-server-dichotomy}

@@

\subsubsection{Stateless}\label{stateless}

@@

\subsubsection{Cacheable}\label{cacheable}

@@

\subsubsection{Uniform Interface}\label{uniform-interface}

@@

\subsubsection{Layered System}\label{layered-system}

@@

Within the MVC pattern, views are the component responsible for the
graphical representation of the application's -- or in this case the
system's -- data
models\cite{krasner1988description}.
Although a view may also be described as ``a visual representation of
models that present a filtered view of the current
state''\cite{osmani2012learning}, the parallels between the MVC pattern and
QuantiTeam's structure become somewhat less applicable. While the
client-side React Native app does of course act as a filtered graphical
representation of the blockchain's models, it also contains its own
local state and therefore deviates from the typical description of an
MVC pattern's view.

@@

The snippet above shows how Flowtype enabled the author to define what
type primitives a \texttt{User} object's fields should adhere to. This
significantly reduced bug frequency, both within the app and the API
itself, by catching inadvertent type coercions or possibly undefined
values before the app's runtime environment was even entered.

Flowtype elevated the ability to define actions and the
expected types of their payloads in a more fine-grained manner compared
to regular JavaScript, by enabling the author to add type constraints to
variables and functions where needed.

@@

\subsubsection{Frequency \& Variability of
Use}\label{frequency-variability-of-use}

@@

\subsubsection{Cross-Platform Potential}\label{cross-platform-potential}

@@

\subsubsection{Synchronous UI Actions}\label{synchronous-ui-actions}

@@

\subsubsection{Asynchronous I/O Actions}\label{asynchronous-io-actions}

@@

Once the required communication abilities for the system had been
established, meaning the author was able to trigger smart contract
events within the blockchain via HTTP requests, the system's API was
constructed and expanded via test-driven development. This meant that
any new functions of the system would be developed in the following
manner\cite{beck2003test}:

@@

============================== Coverage summary ==============================
Statements   : 91.28% ( 429/470 ), 7 ignored
Branches     : 58.02% ( 76/131 ), 7 ignored
Functions    : 98.56% ( 137/139 ), 3 ignored
Lines        : 91.86% ( 429/467 )
================================================================================
\end{lstlisting}

@@

\subsubsection{Verifiability}\label{verifiability}
